---
title: "Assignment 3"
author: "Tobias"
date: "2025-11-09"
output: html_document
---

### 1. Read the data provided in the Excel file and “isolate” the tabulated information into a data.frame class.

```{r}
Holzeinschlag_Hessen <- read.csv(
  file = "C:/Users/Tobias/Documents/GEOmaster/Data_Analysis/data/hessen_holzeinschlag_1997-2014.csv",
  header = TRUE,
  sep = ";",
  skip = 4 ,
  encoding = "latin1")

Holzeinschlag_Hessen <- Holzeinschlag_Hessen[-19, ]

```

### 2. Create a simple visualization which gives a quick, temporally non explicit and descriptive statistical overview of the harvest by tree type and as a total over all trees (i.e. a similar kind of information as provided by the summary function).

```{r}
summary(Holzeinschlag_Hessen)
?color
boxplot(Holzeinschlag_Hessen[,c(3,4,2,5,6,7)],
        main = "Jährlicher Holzeinschlag Hessens 1997-2014", 
        xlab="Baumart ", 
        ylab="Holzeinschlag (m³)", 
        col = "brown4"
        ) 
```

### 3. Create another visualization which - in the same figure panel - shows how each beech harvest over the time span is related to each of the oak, pine, spruce and colored wood harvests in a 2 by 2 grid (i.e. arrange the figures in a 2 columns and 2 rows layout).

```{r}
par(mfrow = c(2,2))

plot(x = Holzeinschlag_Hessen$Buntholz,
     y = Holzeinschlag_Hessen$Buche,
     main = "Bucheneinschlag in Korrelation zu Buntholzeinschlag",
     xlab = "Buntholzeinschlag (m³)",
     ylab = "Buche (m³)",
     cex = 1.5,
     cex.main = 0.8,
     col = "darkgreen",
     pch = 16
     )

plot(x = Holzeinschlag_Hessen$Eiche,
     y = Holzeinschlag_Hessen$Buche,
     main = "Bucheneinschlag in Korrelation zu Eicheneinschlag",
     xlab = "Eiche (m³)",
     ylab = "Buche (m³)",
     cex = 1.5,
     cex.main = 0.8,
     col = "darkgreen",
     pch = 16
     )

plot(x = Holzeinschlag_Hessen$Fichte,
     y = Holzeinschlag_Hessen$Buche,
     main = "Bucheneinschlag in Korrelation zu Fichteneinschlag",
     xlab = "Fichte (m³)",
     ylab = "Buche (m³)",
     cex = 1.5,
     cex.main = 0.8,
     col = "darkgreen",
     pch = 16
     )

plot(x = Holzeinschlag_Hessen$Kiefer,
     y = Holzeinschlag_Hessen$Buche,
     main = "Bucheneinschlag in Korrelation zu Kiefereinschlag",
     xlab = "Kiefer (m³)",
     ylab = "Buche (m³)",
     cex = 1.7,
     cex.main = 0.8,
     col = "darkgreen",
     pch = 16
     )

```

### 4. Include your opinion on what could be the key message of these figures in two sentence max.

```{r}
print("The data suggest that increasing beech harvesting is accompanied by a concurrent rise in the felling of pine and oak, whereas the extraction of spruce remains comparatively stable. Furthermore, I initially hypothesized that an authorized individual had consulted Wikipedia to ascertain that the so-called “Buntholz” does not constitute a valid taxonomic entity—an assumption which, considering the subsequent establishment date of Wikipedia, regrettably proved to be unfounded.")
```
