---
title: "rAssignment2"
author: "Tobias Borzyk"
date: "2025-11-08"
output: html_document
---
### 1.
Implement an if-then-else statement which prints “larger” if the number provided as variable n is larger than three and “equal or smaller” otherwise
- _To check whether the if-statement works, I use the number 5, because I like it._
```{r}
n <- 5
if (n > 3) {
  print("larger")
} else {
  print("equal or smaller")
}
```

### 2.
Extent a copy of the above statement (i.e. copy the entire if-then-else statement and include it a second time in your script in order to preserve both versions) which returns “equal” and “smaller” explicitly in addition to “larger”.
```{r}
if (n > 3) {
  print("larger")
} else if (n < 3) {
  print("smaller")
} else if (n == 3){
  print("equal")
}
```

### 3.
Implement a if-then-else statement which prints “even” if the number provided as variable n is even and which prints “odd” otherwise.
```{r}
if (n %% 2 == 0) {
  print("even")
} else {
  print("odd")
}
```

### 4.
Copy the extended larger/equal/smaller if-then-else statement and include it into a for loop which shows that all three options are actually implemented in a correct manner by iterating over n from a number which is smaller 3, exactly 3 and larger than 3.
```{r}
for (n in 0:5) {
  if (n > 3) {
  print("larger")
} else if (n < 3) {
  print("smaller")
} else if (n == 3){
  print("equal")
}
}
```

### 5.
Extent a copy of the above loop and modify the loop and if-then-else statement in such a way, that the information on “larger” etc. is not printed on the screen but saved within a vector (i.e. a variable which will hold all three statements in the end). Print the content of this vector after the loop.

```{r}
Vektor_Assignment2 <- c()

for (n in 0:5) {
  if (n > 3) {
  Vektor_Assignment2 <- c(Vektor_Assignment2, "larger")
} else if (n < 3) {
  Vektor_Assignment2 <- c(Vektor_Assignment2, "smaller")
} else if (n == 3){
  Vektor_Assignment2 <- c(Vektor_Assignment2, "equal")
}
}

print(Vektor_Assignment2)
```

### 6.
Extent a copy of the above modified loop in such a way, that the results are not saved in a vector but a list. Print the content of this list after the loop.
```{r}
list_Assignment2 <- list()

for (n in 0:5) {
  if (n > 3) {
  list_Assignment2 <- c(list_Assignment2, "larger")
} else if (n < 3) {
  list_Assignment2 <- c(list_Assignment2, "smaller")
} else if (n == 3){
  list_Assignment2 <- c(list_Assignment2, "equal")
}
}

print(list_Assignment2)
```

### 7.
Change the above modified loop in such a way, that the iteration is controlled by a lapply not by a for-loop. Save the returning information from the lapply function in a variable and print the content of this variable after the loop.
```{r}
list_Assignment2_lapply <- lapply(0:5, function(n) {
  if (n > 3) {
    return("larger")
  } else if (n < 3) {
    return("smaller")
  } else if (n == 3){
    return("equal")
  }
})
print(list_Assignment2_lapply)
```

### 8.
Finally change the above variable (i.e. do not modify the loop anymore but just include one more line) in such a way that the content is not printed as a nested list but a vector (i.e. flatten the list).
```{r}
list_Assignment2_lapply <- unlist(lapply(0:5, function(n) {
  if (n > 3) {
    return("larger")
  } else if (n < 3) {
    return("smaller")
  } else if (n == 3) {
    return("equal")
}}))

print(list_Assignment2_lapply)
```

