---
title: "01 Iterate MOHAfun()"
author: "Tobias Borzyk"
date: "2026-01-27"
output: html_document
---
# 01 Iterate MOHAfun() -  Exercises focussing on implementing simple functions and control structures.
1. Implement a function called MOHAfun() which returns “larger” if the number provided as variable x is larger than your matriculation number, “equal” if the number provided as variable x is your matriculation number, and “smaller” if the number provided as variable x is smaller than your matriculation number.
```{r}
MOHAfun <- function(x){
  Matrikelnummer <- 3399925
  if (x > Matrikelnummer){
    return("larger")
  } else if (x == Matrikelnummer){
    return("equal")
  } else if (x < Matrikelnummer){
    return("smaller")
  }
}
```

2. Include MOHAfun() in a for loop which iterates over x in an intervall of 10 integer numbers around your matriculation number. Show the results.

```{r}
numbers_intervall <- 3399920:3399929
ergebnis_numbers_intervall <- list()

for (x in numbers_intervall){
  cat("For x =", x, "the result is:", MOHAfun(x), "\n")
}
```

3. Change the above created loop in such a way that the iteration is controlled by a lapply not by a for-loop. Save the returning information from the lapply function in a variable and print the content of this variable after the loop.
```{r}
ergebnis_numbers_intervall_lapply <- lapply(numbers_intervall, MOHAfun)
print(ergebnis_numbers_intervall_lapply)
```

4. Adjust MOHAfun() by adding y as additional argument instead of your matriculation number in order to allow x to be compared with any input.
```{r}
MOHAfun <- function(x,y){
  if (x > y){
    return("larger")
  } else if (x == y){
    return("equal")
  } else if (x < y){
    return("smaller")
  }
}
```

